---
output:
  pdf_document: default
---
## Problem 5: Computational Finance - Modelling Stock prices

Following piece of code download the prices of TCS since 2007

```{r}
library(quantmod)
getSymbols('TCS.NS')
tail(TCS.NS)
```
Plot the adjusted close prices of TCS
```{r}
plot(TCS.NS$TCS.NS.Adjusted)
```

**Download the data of market index Nifty50**. The Nifty 50 index indicates how the over all market has done over the similar period.
```{r}
getSymbols('^NSEI')
tail(NSEI)
```
Plot the adjusted close value of Nifty50
```{r}
plot(NSEI$NSEI.Adjusted)
```


### Log-Return 
We calculate the daily log-return, where log-return is defined as
$$
r_t=\log(P_t)-\log(P_{t-1})=\Delta \log(P_t),
$$
where $P_t$ is the closing price of the stock on $t^{th}$ day.

```{r}
TCS_rt = diff(log(TCS.NS$TCS.NS.Adjusted))
Nifty_rt = diff(log(NSEI$NSEI.Adjusted))
retrn = cbind.xts(TCS_rt,Nifty_rt) 
retrn = na.omit(data.frame(retrn))

plot(retrn$NSEI.Adjusted,retrn$TCS.NS.Adjusted
     ,pch=20
     ,xlab='Market Return'
     ,ylab='TCS Return'
     ,xlim=c(-0.18,0.18)
     ,ylim=c(-0.18,0.18))
grid(col='grey',lty=1)
```

+ Consider the following model:

$$
r_{t}^{TCS}=\alpha + \beta r_{t}^{Nifty} + \varepsilon,
$$
where $\mathbb{E}(\varepsilon)=0$ and $\mathbb{V}ar(\varepsilon)=\sigma^2$.

1. Estimate the parameters of the models $\theta=(\alpha,\beta,\sigma)$ using the method of moments type plug-in estimator discussed in the class.

## Solution:
```{r}
mu_y = mean(retrn$TCS.NS.Adjusted)
mu_x = mean(retrn$NSEI.Adjusted)
sigma_y = sd(retrn$TCS.NS.Adjusted)
sigma_x = sd(retrn$NSEI.Adjusted)
rho = cor(retrn$NSEI.Adjusted, retrn$TCS.NS.Adjusted)

alpha_0 = mu_y - rho * mu_x * (sigma_x/sigma_y)
beta_0 = rho *  (sigma_x/sigma_y)

retrn_mom = retrn
retrn_mom$mom_est_TCS.NS.Adjusted = alpha_0 + beta_0 * retrn_mom$NSEI.Adjusted

epsilon = retrn_mom$TCS.NS.Adjusted-retrn_mom$mom_est_TCS.NS.Adjusted
sigma_0 = sd(epsilon)

theta_0 = c(alpha_0, beta_0, sigma_0)

print( theta_0)
```

2. Estimate the parameters using the `lm` built-in function of `R`. Note that `lm` using the OLS method.

## Solution:
```{r}
lm_model = lm(TCS.NS.Adjusted~NSEI.Adjusted, data = retrn)

co = lm_model$coefficients

alpha_1 = matrix(co)[1,1]
beta_1 = matrix(co)[2,1]
sigma_1 = sd(lm_model$residuals)

theta_1 = c(alpha_1, beta_1, sigma_1)

print(theta_1)
```


3. Fill-up the following table

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |                   |
$\beta$    |                   |
$\sigma$   |                   |

## Solution:

Parameters | Method of Moments | OLS
-----------|-------------------|-----
$\alpha$   |  $0.0005848199$   |$0.0004611208$
$\beta$    |  $0.3904739457$   |$0.7436970826$
$\sigma$   |  $0.0169171740$   |$0.0161865264$

4. If the current value of Nifty is 18000 and it goes up to 18200. The current value of TCS is Rs. 3200/-. How much you can expect TCS price to go up?

## Solution:

```{r}
prediction_mom = function(Nifty_initial_value, Nifty_final_value, TCS_initial_value){
  beta = theta_0[2]
  alpha = theta_0[1]
  x = log(Nifty_final_value) - log(Nifty_initial_value)
  y = alpha + beta*x
  y1 = log(TCS_initial_value) + y
  y2 = exp(y1)
  return(y2)
}

prediction_ols = function(Nifty_initial_value, Nifty_final_value, TCS_initial_value){
  beta = theta_1[2]
  alpha = theta_1[1]
  x = log(Nifty_final_value) - log(Nifty_initial_value)
  y = alpha + beta*x
  y1 = log(TCS_initial_value) + y
  y2 = exp(y1)
  return(y2)
}
```

By the Method of Moments type method, we can say that we can expect TCS value to go up by
```{r}
Nifty_initial_value = 18000
Nifty_final_value = 18200
TCS_initial_value = 3200

TCS_final_value = prediction_mom(Nifty_initial_value = Nifty_initial_value,
                                 Nifty_final_value = Nifty_final_value,
                                 TCS_initial_value = TCS_initial_value)

TCS_final_value - TCS_initial_value
```

By the OLS method, we can say that we can expect TCS value to go up by
```{r}
TCS_final_value = prediction_ols(Nifty_initial_value = Nifty_initial_value,
                                 Nifty_final_value = Nifty_final_value,
                                 TCS_initial_value = TCS_initial_value)

TCS_final_value - TCS_initial_value
```

Thus we can expect TCS price to go up by some value around Rs. 15.72 and Rs. 27.89.